### 2017-04-08 SpringRateProject/MSXII Supsension Parameters.tex
Fucked up the units
Fixed mistake made when finding applied loads on the vehicle (didn’t
convert kN to N in calculations). Updated report and matlab script
accordingly. Reorganized file structure

\subsection{Selecting Suspension Parameters}
\label{sec:paramSelection}
The spring rates are calculated based on the travel of the wheel and the maximum vertical loading of the wheel due to normal force from the road. Assuming a smooth driving surface, the largest normal force acting on the tire will occur while braking in a turn. To approximate this loading condition, the a superposition of two static half car models is used; one for cornering and one for hard braking.  
\subsubsection{Hard Braking}\begin{figure}[h!]
	\centering
	\includegraphics[width=.5\textwidth]{./LaTex/brakingFBD.jpg}
	\caption{Free Body Diagram of a Car Braking}
	\label{fig:barkingFBD}
\end{figure}
Figure \ref{fig:barkingFBD} shows a free body diagram of a half car model undergoing braking. MSXII will only have brake callipers on the front wheels however regenerative braking from the hub motors will also provide a braking force at the rear contact patches. Assuming a generous coefficient of static friction, $\mu _s$, of 1.0 and the extreme case where the both the font and rear tires are about to begin sliding, the combined braking force can be expressed as: 
\begin{equation}
	F_{braking} = F_{Nf} + F_{Nr} = \mu _s F_{N_{net}} =  \mu _s Mg = (1.0)(550kg)\left(9.81\frac{m}{s^2}\right) = 5.40kN
\end{equation}
From Figure \ref{fig:barkingFBD}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
\begin{equation}
	aF_{Nf} = hF_{braking} + bF_{Nr}
\end{equation}
\begin{equation}
	F_{Nf} + F_{Nr} = Mg
\end{equation}
Solving for the front normal force: 
\begin{equation}
\begin{split}
	aF_{Nf} &= hF_{braking} + b(Mg - F_{Nr})\\
	F_{Nf} &= \frac{hF_{braking} + bMg}{a+b} = \frac{(0.52m)(5.40kN)+(1.28m)(550kg)\left(9.81\frac{m}{s^2}\right)}{(1.32m)+(1.28m)} = 2.66kN
\end{split}
\end{equation}
Note that 2.66kN is for both front wheels. Therefore there is approximately \textbf{1.33kN} of normal force on each front tire during hard braking.

@@

\noindent For the half car model shown if Figure \ref{fig:steerFBD}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
\begin{equation}
	\frac{track}{2}F_{Nf} = hF_{steer} + \frac{track}{2}F_{Nr}
\end{equation}
\begin{equation}
	F_{No} + F_{Ni} = Mg
\end{equation}
Solving for the outside wheel normal force: 
\begin{equation}
\begin{split}
	\frac{track}{2}F_{No} &= hF_{steer} + \frac{track}{2}(Mg - F_{Ni})\\
	F_{No} &= \frac{hF_{steer} + \frac{track}{2}Mg}{track} = \frac{(0.52m)(2.28kN)+(0.80m)(550kg)\left(9.81\frac{m}{s^2}\right)}{1.60m} = 2.70kN
\end{split}
\end{equation}
Note that the 2.70kN above is for both front and rear outside wheels. Assuming the force is evenly split between the front and rear, there is approximately \textbf{1.35kN} of normal force on each outside tire during maximum cornering. 
\subsubsection{Spring Rate Selection}
Adding the maximum normal forces from cornering and braking results in 2.68kN. The maximum travel of the suspension for MSXII in compression has been set to 45mm. It is desired that under the combined loading condition of braking and cornering, the suspension compresses to a maximum of 70\% of the travel. As such, the desired spring rate is found to be: 
\begin{equation}
	k = \frac{force}{displacement} = \frac{2.68N}{(70\%)(0.045m)} = 85.0kN/m
\end{equation}
Note that the above rate assumes a vertical spring. The vertical spring rate can be mapped to an equivalent angled spring rate by dividing by the cosine of the angle that the coilover makes with the vertical direction: 
\begin{equation}
	\begin{split}
		k_{front} &= \frac{85.0}{cos(35)} = 110.6kN \\
		k_{rear} &= {85.0}{cos(30)} = 98.1kN
	\end{split}
\end{equation}
MSXII will be using Ohlins TTX25 MkII coilovers for both the front and rear. These coilovers can be configured to have spring rates from 150 lb/in to 650 lb/in in increments of 50 lb/in. 1 lb/in is approximately 175 N/m therefore 550 lb/in coils should be selected for the front and 650 lb/in coils for the rear.

\subsubsection{Damping Coefficient Selection}
Ohlins has published damper curves for their TTX25 MkII's however real world damping rates are non constant. For this report, the damping coefficient will be idealized as constant and will be selected such that the damping ratio, $\zeta$, is 0.5. This is typical of high performance consumer vehicles \cite{dampers}. Given an equivalent vertical spring rate or 85.0kN/m, a damping ratio of 0.5 maps to a vertical damping coefficient of \textbf(4.83kNs/m).

@@

\subsubsection{Solution}
Based on the idealize parameters proposed in Section \ref{sec:paramSelection} , the natural frequency and damping ratio are found as follows: 
\begin{equation}
	\begin{split}
	\omega _{nx} &= \sqrt{\frac{2k}{M}} = \sqrt{\frac{(2)(85000N/m)}{550kg}} = 17.58rad/s = 2.80Hz \\
	\omega _{n\Theta} &= \sqrt{\frac{2a^2k}{I}} = \sqrt{\frac{(2)(1.3^2m^2)(85000N)}{550kgm^2}} = 22.86rad/s = 3.64Hz 
	\end{split}  
\end{equation}
\begin{equation}
	\begin{split}
		c_{eqx} &= 2c = (2)(4830Ns/m) = 9660Ns/m \\ 
		c_{rx} &= 2M\omega _n = (2)(550kg)(17.58rad/s) = 19340Ns/m \\ 
		\zeta _x& = \frac{c_{eqx}}{c_r} = \frac{9660Ns/m}{19340Ns/m} = 0.50\\
		c_{eq\Theta} &= 2a^2c =  2(1.3^2m^2)(4830Ns/m) = 16330Nms\\
		c_{r\Theta} &= 2m_{eq}\omega _n = (2)(550kgm^2)(22.86rad/s) = 25150Nms\\ 
		\zeta _\Theta &= \frac{c_{eq\Theta}}{c_r} = \frac{16330Nms}{ 25150Nms} = 0.65
	\end{split}  
\end{equation}
Using the MATLAB script shown in Appendix \ref{app:simple}, values for the parameters $I$, $M$, $k$, $c$, and $a$ are substituted into Equations \ref{eq:eomx} and \ref{eq:eomtheta} and the Laplace transform is taken. Solving for $X(s)$ and $\Theta(s)$ and taking the inverse Laplace transform yeilds:
\begin{equation}
		\begin{split}
		x(t) = &(3.12\cdot 10^{-5})\, \sin\!(69.7\, t) - 0.00201\, \cos\!(69.7\, t) + 1.0\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, ((5.91\cdot 10^{-4})\, \cos\!(69.7\, t) \\
		&- 0.00192\, \sin\!(69.7\, t)) + \mathrm{e}^{- 8.78\, t}\, (0.00201\, \cos\!(15.2\, t) + 0.00102\, \sin\!(15.2\, t) \\
		&- \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, (0.00396\, \cos\!(15.2\, t) - 0.00489\, \sin\!(15.2\, t)))
		\end{split}
\end{equation}

>>>>

\subsection{Selecting Suspension Parameters}
\label{sec:paramSelection}
The spring rates are calculated based on the travel of the wheel and the maximum vertical loading of the wheel due to normal force from the road. Assuming a smooth driving surface, the largest normal force acting on the tire will occur while braking in a turn. To approximate this loading condition, the a superposition of two static half car models is used; one for cornering and one for hard braking.  
\subsubsection{Hard Braking}\begin{figure}[h!]
	\centering
	\includegraphics[width=.5\textwidth]{./LaTex/brakingFBD.jpg}
	\caption{Free Body Diagram of a Car Braking}
	\label{fig:barkingFBD}
\end{figure}
Figure \ref{fig:barkingFBD} shows a free body diagram of a half car model undergoing braking. MSXII will only have brake callipers on the front wheels however regenerative braking from the hub motors will also provide a braking force at the rear contact patches. Assuming a generous coefficient of static friction, $\mu _s$, of 1.0 and the extreme case where the both the font and rear tires are about to begin sliding, the combined braking force can be expressed as: 
\begin{equation}
	F_{braking} = F_{Nf} + F_{Nr} = \mu _s F_{N_{net}} =  \mu _s Mg = (1.0)(550kg)\left(9.81\frac{m}{s^2}\right) = 5.40kN
\end{equation}
From Figure \ref{fig:barkingFBD}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
\begin{equation}
	aF_{Nf} = hF_{braking} + bF_{Nr}
\end{equation}
\begin{equation}
	F_{Nf} + F_{Nr} = Mg
\end{equation}
Solving for the front normal force: 
\begin{equation}
\begin{split}
	aF_{Nf} &= hF_{braking} + b(Mg - F_{Nf})\\
	F_{Nf} &= \frac{hF_{braking} + bMg}{a+b} = \frac{(0.52m)(5.40kN)+(1.28m)(550kg)\left(9.81\frac{m}{s^2}\right)}{(1.32m)+(1.28m)} = 3.74kN
\end{split}
\end{equation}
Note that 3.74kN is for both front wheels. Therefore there is approximately \textbf{1.87kN} of normal force on each front tire during hard braking.

@@

\noindent For the half car model shown if Figure \ref{fig:steerFBD}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
\begin{equation}
	\frac{track}{2}F_{No} = hF_{steer} + \frac{track}{2}F_{Ni}
\end{equation}
\begin{equation}
	F_{No} + F_{Ni} = Mg
\end{equation}
Solving for the outside wheel normal force: 
\begin{equation}
\begin{split}
	\frac{track}{2}F_{No} &= hF_{steer} + \frac{track}{2}(Mg - F_{Ni})\\
	F_{No} &= \frac{hF_{steer} + \frac{track}{2}Mg}{track} = \frac{(0.52m)(2.28kN)+(0.80m)(550kg)\left(9.81\frac{m}{s^2}\right)}{1.60m} = 3.44kN
\end{split}
\end{equation}
Note that the 3.44kN above is for both front and rear outside wheels. Assuming the force is evenly split between the front and rear, there is approximately \textbf{1.72kN} of normal force on each outside tire during maximum cornering. 
\subsubsection{Spring Rate Selection}
Adding the maximum normal forces from cornering and braking results in 3.59kN. The maximum travel of the suspension for MSXII in compression has been set to 45mm. It is desired that under the combined loading condition of braking and cornering, the suspension compresses to a maximum of 90\% of the travel. As such, the desired spring rate is found to be: 
\begin{equation}
	k = \frac{force}{displacement} = \frac{3.59kN}{(70\%)(0.045m)} = 88.6kN/m
\end{equation}
Note that the above rate assumes a vertical spring. The vertical spring rate can be mapped to an equivalent angled spring rate by dividing by the cosine of the angle that the coilover makes with the vertical direction: 
\begin{equation}
	\begin{split}
		k_{front} &= \frac{88.6}{cos(35)} = 108.2kN \\
		k_{rear} &= {88.6}{cos(30)} = 102.3kN
	\end{split}
\end{equation}
MSXII will be using Ohlins TTX25 MkII coilovers for both the front and rear. These coilovers can be configured to have spring rates from 150 lb/in to 650 lb/in in increments of 50 lb/in. 1 lb/in is approximately 175 N/m therefore 600 lb/in coils should be selected for the front and 600 lb/in coils for the rear.

\subsubsection{Damping Coefficient Selection}
Ohlins has published damper curves for their TTX25 MkII's however real world damping rates are non constant. For this report, the damping coefficient will be idealized as constant and will be selected such that the damping ratio, $\zeta$, is 0.5. This is typical of high performance consumer vehicles \cite{dampers}. Given an equivalent vertical spring rate or 88.6/m, a damping ratio of 0.5 maps to a vertical damping coefficient of \textbf(4.83kNs/m).

@@

\subsubsection{Solution}
Based on the idealize parameters proposed in Section \ref{sec:paramSelection} , the natural frequency and damping ratio are found as follows: 
\begin{equation}
	\begin{split}
	\omega _{nx} &= \sqrt{\frac{2k}{M}} = \sqrt{\frac{(2)(88600N/m)}{550kg}} = 17.95rad/s = 2.86Hz \\
	\omega _{n\Theta} &= \sqrt{\frac{2a^2k}{I}} = \sqrt{\frac{(2)(1.3^2m^2)(88600N)}{550kgm^2}} = 23.33rad/s = 3.71Hz 
	\end{split}  
\end{equation}
\begin{equation}
	\begin{split}
		c_{eqx} &= 2c = (2)(4830Ns/m) = 9660Ns/m \\ 
		c_{rx} &= 2M\omega _n = (2)(550kg)(17.75rad/s) = 19530Ns/m \\ 
		\zeta _x& = \frac{c_{eqx}}{c_r} = \frac{9660Ns/m}{19530Ns/m} = 0.49\\
		c_{eq\Theta} &= 2a^2c =  2(1.3^2m^2)(4830Ns/m) = 16330Nms\\
		c_{r\Theta} &= 2m_{eq}\omega _n = (2)(550kgm^2)(23.33rad/s) = 25660Nms\\ 
		\zeta _\Theta &= \frac{c_{eq\Theta}}{c_r} = \frac{16330Nms}{ 25660Nms} = 0.64
	\end{split}  
\end{equation}
Using the MATLAB script shown in Appendix \ref{app:simple}, values for the parameters $I$, $M$, $k$, $c$, and $a$ are substituted into Equations \ref{eq:eomx} and \ref{eq:eomtheta} and the Laplace transform is taken. Solving for $X(s)$ and $\Theta(s)$ and taking the inverse Laplace transform yeilds:
\begin{equation}
		\begin{split}
		x(t) = &(3.12\cdot 10^{-5})\, \sin\!(69.7\, t) - 0.00201\, \cos\!(69.7\, t) + 1.0\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, ((5.91\cdot 10^{-4})\, \cos\!(69.7\, t) \\
		&- 0.00192\, \sin\!(69.7\, t)) + \mathrm{e}^{- 8.78\, t}\, (0.00201\, \cos\!(15.2\, t) + 0.00102\, \sin\!(15.2\, t) \\
		&- \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, (0.00396\, \cos\!(15.2\, t) - 0.00489\, \sin\!(15.2\, t)))
		\end{split}
\end{equation}
